---
title: "Homework5"
author: "Laurenz Ohnemüller"
date: "10/7/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Question 1

Volume of the unit ball:  $v_n = \frac{\pi^{\frac{n}{2}}}{\Gamma(\frac{n}{2}+1)}$ \newline
Gamma function: $\Gamma(a) = \int_0^\infty x^ae^{-x}\frac{dx}{x}$

(a)\newline
First of all, we want to compute the probability that a point is in the unit ball in the n-dimensional space. 
Let X denote that $(U_1,U_2,...,U_n)$ is in the unit ball in $R^n$ then:

$$P(X_n) = \frac{volume\_of\_ball}{area\_of\_square}$$
We can calculate this by dividing the volume of the unit ball $v_n$ by the area of the square of dimension $n$. The volume for the unit ball in the nth dimension is already given, so we only need to figure out the area of the square. \newline
Since $(U_1,U_2,...,U_n) \sim Unif(-1,1)$, we can imagine the square going from -1 to 1, which means that one side has the length of 2. Now we can just take it to the power of $n$, to find the area of the $nth$ dimension. So:

$$P(X_n) = \frac{v_n}{2^n} = \frac{\frac{\pi^{\frac{n}{2}}}{\Gamma(\frac{n}{2}+1)}}{2^n} =  \frac{\pi^{\frac{n}{2}}}{\Gamma(\frac{n}{2}+1){2^n}}$$

Finally, we just have to insert a specific dimension and we get the probability that this particular point is in the unit ball.


(b)

```{r Question 1b}
# I assumed we are allowed to use this library, since we also used it multiple times in the Exercises
library("tibble")
trials = 10
p = tibble(n=1:trials,
           (pi^(n/2))/(gamma(n/2+1)*2^n)
)
plot(p, ylab = "probability", xlab = "trials", main="Plotted Probabilites")
```

(c)
We can see the condition $|U_j| > c$ as a Bernoulli experiment, either this is true ("success") or it is not true ("failure"). Having $n$ of these Bernoulli trials would speak for a Binomial distribution, and to that we also are sure about the independence of $U_j$, which insures us that the probability p does not change. \newline
So: 
$$X_n \sim Bin(n,p) $$ \newline
$$X_n = \binom{n}{k}p^k(1-p)^{n-k}$$ \newline

(d)
Imagine a circle drawn inside a square - to make it more simple, we can do that in the third dimension, so with a ball inside a cube. We now that each edge has the length 2, since one goes from -1 to 1, while the ball also surrounds the origin (0,0). We want to find the probability that a random point ($U_1, U_2, U_3$) is inside this unit ball.



Question 2

The definition of the PMF of a joint distribution is:

$$p_{x,y} = P(X=x, Y=y)$$

(a)
In this question, we are supposed to find the joint PMF of X, Y and N. 
$$P(X=x, Y=y, N=n)$$
Note that x, y and n are nonnegative integers. Also note that since x describes the number of times the traveler is lost and asks for direction and y the number of times he does not ask for direction, while n describes the amount of times he is lost in generel. This sum $x+y=n$ must hold, otherwise the PMF is 0. \newline

Now we apply the Law of Total Probability and condition on N: 

$$P(X=x, Y=y, N=n) = \sum_{n=0}^\infty P(X=x, Y=y | N=n) P(N=n) = P(X=x, Y=y | N=n) P(N=n)$$
Hereby, we can take the N out, since we know that it holds in the case where $x+y=n$, which means we can replace it by $N = X+Y$.

$$P(X=x, Y=y , N=n) = P(X=x, Y=y, X+Y = n) = P(X=x, Y=y, X = n-Y) =P(X=x, Y=y, X = n-y)$$
We can observe that $X = n-y$ is redundant information, since it the exact same thing as $X = x$, so we end up here: 

$$P(X=x, Y=y)$$
From the chicken and egg story (and also proven in (c)), we know that X and Y are independet and Possion distributed:\newline
$X \sim Pois(\lambda p)$ \newline
$Y \sim Pois(\lambda q)$, whereby q = 1 - p \newline
, which brings us to this:
$$P(X=x, Y=y) = P(X=x)P(Y=y) = \frac{e^{-\lambda p}(\lambda p)^x}{x!} \frac{e^{-\lambda q}(\lambda q)^y}{y!}$$
Lastly, we want to check wheather X, Y and   are independent or not.
For independence this condition must hold: $P(X=x, Y=y , N=n) = P(X=x)P(Y=y)P(N=n)$
We can already see that this will not hold, since we just take the calculated value from above and multiply it bei $P(N=n)$, which is also poisson distributed $N \sim Pois(\lambda)$ and it won't equal each other.
Summarizing, (X,Y,N) are dependent.
  

(b)

In the second part we want to find the joint PMF of X and N, which we can express in the following way using again the Law of Total Probability: 

$$P(X=x, N=n) = P(X=x|N=n)P(N=n)$$
Here, we plug in the defintions of the Binomial (for $P(X=x|N=n)$)and Possion distribution (for $P(N=n)$):


$$= \binom{n}{x} p^x q^{n-x} * \frac{e^{-\lambda}\lambda^n}{n!}$$    

To check wheather they are dependent or independent, we use this condition of independence: 

$$P(X=x, N=n) = P(X=x)P(N=n)$$
We first calculate $P(X=x)$, where the random variable has the folllowing distribution $X \sim Pois(\lambda p)$:

$$P(X=x) = \frac{e^{-\lambda p}*(\lambda p)^x}{x!}$$
and since $P(N=n)$ is normally Poission distributed, we get: 

$$P(X=x)P(N=n) = \frac{e^{-\lambda p}*(\lambda p)^x}{x!}* \frac{e^{-\lambda}*\lambda^n}{n!} \neq P(X=x, N=n)$$
This means that N and X are indeed dependent.

(c)

Finally, we are supposed to find: 

$$P(X=x, Y=y)$$
, where we condition on N and apply the Law of total probability (LOTP), we assume we know the amount of times our traveler was lost.


$$P(X=x, Y=y) = \sum_{n=0}^\infty P(X = x, Y = y |N = n)P(N=n)$$
Since $n$ describes the amount of times the traveler gets lost, while $x$, as well as $y$, describe if he asked for directions or not in the case that he is lost, we can say $x+y=n$, otherwise the probability is zero. \newline
Knowing this, we can transform our equation to: 

$$P(X = x, Y = y| N = x+y)P(N=x+y)$$
Now we can observe that that $Y=y$ is redundant, since it is exactly the same event than $X=x$. \newline
So we are left with: 

$$P(X = x,| N = x+y)P(N=x+y)$$
, which is easy to compute using the definitions of the Binomial and Possion distribution. \newline 

Binomial distribution: $\frac{n!}{k!(n-k)!}p^kq^{n-k}$, where n = x+y, and k = x\newline  
\newline
Possion distribution: $\frac{e^{-\lambda}\lambda^k}{k!}$, where k = x+y \newline

$$= \frac{(x+y)!}{x!y!}p^xq^y*\frac{e^{-\lambda}\lambda^{x+y}}{(x+y)!}$$
Here we can reduce and reformate the equation, sucht that: 

$$= e^{-\lambda p}\frac{(\lambda p)^x}{x!}*e^{-\lambda q}\frac{(\lambda q)^y}{y!}$$
On this equation, we can observe that X and Y are independent, since the joint PMF is the product of $X \sim Pois(\lambda p)$ and $X \sim Pois(\lambda q)$, which makes this condition hold:  $P(X=x, Y=y) =P(X=x)P(Y=y)$.





Question 3

The joint PDF of X and Y: 

$$f_{X,Y}(x,y) = c e^{-\frac{x^2}{2}}e^{-\frac{y^2}{2}}$$

(a)

To be a valid PDF $f_{X,Y}(x,y)$ must satisfy two criteria: 

  (1) Nonnegative: $f_{X,Y}(x,y) \geq 0$; \newline
  (2) Integrates to 1: $\int_{-\infty}^\infty \int_{-\infty}^\infty f_{X,Y}(x,y) dxdy= 1$ \newline

This allows us directly to define c as $c \geq 0$, since otherwise (1) would be violated.
To find the actual c, we have to solve the integral:

$$ \int_{-\infty}^\infty \int_{-\infty}^\infty c e^{-\frac{x^2}{2}}e^{-\frac{y^2}{2}} dxdy = 1$$ 
\newline
c is constant, which means that we can pull it before the integrals: 

$$c\int_{-\infty}^\infty \int_{-\infty}^\infty e^{-\frac{x^2}{2}}e^{-\frac{y^2}{2}} dxdy = 1$$
Now, we will just focus on solving this integral - without thinking about the 1 or the c:

$$\int_{-\infty}^\infty \int_{-\infty}^\infty e^{-\frac{x^2+y^2}{2}} dxdy$$
Here we make use of the polar coordinates (A.7.2 in appendix in our Probability book). \newline
$x=r*cos(\theta), y= r*sin(\theta)$, where r is the distance from (x,y) to the origin and $\theta \in [0,2\pi]$ is the angle. Hereby, $dxdy$ becomes $rdrd\theta$ and $r = \sqrt{x^2+y^2}$ (so r must be $r\geq0$). This changes also the limits, as already explained $\theta$ goes from $0$ to $2\pi$, while r can take on the values form $0$ to $\infty$ and it follows:
                          
$$\int_0^{2\pi} \int_{0}^\infty e^{-\frac{r^2}{2}} rdrd\theta$$
This integral is now solvable, using substitution $u= \frac{r^2}{2}; du = rdr$:
\begin{align}
\int_0^{2\pi}(\int_{0}^\infty e^{-u} du)d\theta = \int_0^{2\pi}(-e^{-u} |_0^\infty)d\theta = \int_0^{2\pi}1 d\theta = 2\pi 
\end{align}

Finally, we can plug this result into our equation from earlier: 

$$c*2\pi = 1 \rightarrow c=\frac{1}{2\pi}$$
(b)

Since our distribution is continous, we need the definition of the Marginal PDF of X: 

$$f_X(x) = \int_{-\infty}^\infty f_{X,Y}(x,y)dy$$
From here on we can continue our work similarly to the previous question:

$$f_X(x) = \int_{-\infty}^\infty c e^{-\frac{x^2}{2}}e^{-\frac{y^2}{2}} dy = c e^{-\frac{x^2}{2}} \int_{-\infty}^\infty e^{-\frac{y^2}{2}} dy$$
\newline
To figure out $\int_{-\infty}^\infty e^{-\frac{y^2}{2}} dy$, we can just look at (a) again, and in particular we can focus on equation (1), where we proved that: 

$$\int_{-\infty}^\infty \int_{-\infty}^\infty e^{-\frac{x^2}{2}}e^{-\frac{y^2}{2}} dxdy = 2\pi$$
We can split that integral apart and summarize it as, since $x$ and $y$ are just different names for the "dummy" variable $z$: 

$$\int_{-\infty}^\infty e^{-\frac{x^2}{2}} dx \int_{-\infty}^\infty e^{-\frac{y^2}{2}} dy = (\int_{-\infty}^\infty e^{-\frac{y^2}{2}} dy)^2 = 2\pi$$
Out of this equality, we find the value for our integral: 

$$\int_{-\infty}^\infty e^{-\frac{y^2}{2}} dy = \sqrt {2\pi}$$
\newline
Now lets put everything together again and plug in the values for the integral and c:

$$f_X(x) = c e^{-\frac{x^2}{2}} \int_{-\infty}^\infty e^{-\frac{y^2}{2}} dy = \frac{1}{2\pi} e^{-\frac{x^2}{2}} *\sqrt{2\pi} = \frac{1}{\sqrt{2\pi}}e^{-\frac{x^2}{2}}$$
On the exact same way, we can also find out $f_Y(y)$, which will be - because of symmetry - the same as $f_Y(y)$: 
$$f_Y(y) = \frac{1}{\sqrt{2\pi}}e^{-\frac{y^2}{2}}$$
This is by the way the standard normal distribution.
\newline
To check wheather the X and Y are independent, we examine this condition:

$$f_{X,Y}(x,y) = f_X(x)f_Y(y)$$
Let us now insert our solutions and check the condition:

$$\frac{1}{2\pi} e^{-\frac{x^2}{2}}e^{-\frac{y^2}{2}} \stackrel{?}{=} \frac{1}{\sqrt{2\pi}}e^{-\frac{x^2}{2}}*\frac{1}{\sqrt{2\pi}}e^{-\frac{y^2}{2}} $$
This is actually true, which confirms the condition and makes the X and Y independent.
\newline 
(c) \newline
First of all, we know that X and Y are both marginally $\mathcal{N}(0,1)$.
For $(X,Y)$ to be a Bivariate Normal, $aX+bY$ has to have a normal distribution for $a,b \in \mathbb{R}$. \newline
This is certainly true,the sum of independent Normals is also Normal. Any value for $a$ or $b$, won't change that. 
Also for $a=b=0$, where $aX+bY = 0$ is still normally distributed with mean and variance 0. 

